Software Clusterings with Vector Semantics and the Call Graph

### 2019-07-04 poster/poster.tex
Add slides with makefile

	\begin{itemize}
    \item The architecture of a software system is the most fundamental realization of it
    \item When there is no specific definition of it, we can attempt to recover it
    \item One particular problem is the \textbf{clustering of its components into modules} 
    \item Many methods exist in literature such as agglomerative clustering baselines as well as more sophisticated algorithms, like LIMBO and ACDC. 
\end{itemize}

@@

\begin{enumerate}
    \item \textbf{Provide} a method for software clusterings through \textbf{vector semantics} and the \textbf{call graph} to determine software's modules without knowledge of its structure
    \item Empirically validate our method performance on the \textbf{Linux Kernel Codebase}

    \item Compare it against state-of-the-art methods (ACDC \cite{acdc}, LIMBO \cite{LIMBO}) and agglomerative clustering methods (agglomerative clustering \cite{average, complete, ward}) 
\end{enumerate}

@@

\item Andritsos and Tzerpos in reference \cite{LIMBO} present an Information-Theoretic 
approach of \textbf{sca} by developing \textbf{LIMBO} which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \textbf{Agglomerative Information Bottleneck} algorithm.
\end{itemize}

@@

\item The edges have the normalized cosine similarities as weights $w(i,j) = (1 + \cos(\vec x_i, \vec x_j)) / 2$
\end{enumerate}

\heading{Obtaining a clustering}
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
The Louvain Method is a greedy optimization method that maximizes the modularity function of a graph $H(V, E)$

@@

\item it is a large and complex system with 27 years of continuous development 
\item it spans $\sim$20.3 million lines of source code at the time of writing
\item  it is easy to establish a ground truth due to its clear structure and construct test-cases for evaluations.

\item it is used in related studies \cite{acdc, evaluation} as a reference system for evaluation

@@

\item Compare our method against agglomerative clustering baselines (average, complete, and Ward linkage) and state-of-the-art methods (LIMBO and ACDC)
\item Use the MoJo distance metric to calculate the distance between the proposed clustering and ground truth
\item Evaluate clusterings \textbf{stability}, \textbf{authoritativeness} and \textbf{extremity}

@@

    \begin{table}
      \centering
      \small
\begin{tabular}{lrrrrrrr}
    \toprule
    \textbf{Algorithm} & \textbf{Feat. Dim}  & $n_c$ & \textbf{Range} & $\bar x$ & $\sigma$ & \textbf{Median} & \textbf{MoJo D.} \\
    \midrule
    \textbf{ACDC}  & -- & 9055 & 1 -- 4245 & 5 & 46 & 2 & 33694\\
    Average Linkage  & \textbf{300} & \textbf{21} & 1--3406 & 163 & 725 & 1 & 2092 \\
    Complete Linkage  & \textbf{300} & \textbf{21} & 1--1529 & 163 & 412 & 19 & 1710 \\
    \textbf{LIMBO}  ($B=100, \; S = \infty$) & 12317 &\textbf{21} & 50--1810 & 163 & 375 & 50  & 1482 \\

    Ward Linkage & \textbf{300} & \textbf{21} & 21--948 & 163 & 223 & 70 & 1138 \\

    \textbf{SADE} & \textbf{300} & 10 ($\pm$ 2)  & 2 ($\pm$ 0) -132 ($\pm$ 13) & 64 ($\pm$ 4) & 40 ($\pm$ 4) & 65 ($\pm$ 10) & 243 ($\pm$ 1)  \\
    \textbf{SADE} (Directed) & \textbf{300} & 5 ($\pm$ 2) & 1 ($\pm$ 1) - 614 ($\pm$ 1) & 141 ($\pm$ 39) & 253 ($\pm$ 25) & 2 ($\pm$ 0.3)  & 237 ($\pm$ 2) \\
    \midrule
    Ground Truth & -- & 21 & 1--1348 & 163 & 341 & 11.0 & -- \\
    \bottomrule
  \end{tabular}

@@

	\begin{itemize}
		\justifying
    \item \textbf{Surpass} all clustering algorithms in terms of MoJo Distance Metric
    \item Production of \textbf{balanced clusterings}
    \item Production of \textbf{stable clusterings}
    \item Results were produced without knowing the number of clusters of the ground truth a priori
    \item Provide a \textbf{simplistic approach} to software clustering combining vector semantics and the call graph
	\end{itemize}

@@

	\begin{alertblock}{Conclusions}
	\begin{enumerate}
	\justifying
    \item Use of vector semantics and the call graph to produce meaningful clusterings
    \item Performing our study on a very large system (Linux) gives us further insight on the nature of software itself
    \item Outperform state-of-the-art and baseline methods in terms of authoritativeness and extremity
    \item Produce stable and balanced clusterings 
\end{enumerate}

>>>>

	\begin{itemize}
    \item The architecture of a software system is the most fundamental realization of it.
    \item When there is no specific definition of it, we can attempt to recover it.
    \item One particular problem is the \textbf{clustering of its components into modules}. 
    \item Many methods exist in literature such as agglomerative clustering baselines as well as more sophisticated algorithms, like LIMBO and ACDC. 
\end{itemize}

@@

\begin{enumerate}
    \item \textbf{Provide} a method for software clusterings through \textbf{vector semantics} and the \textbf{call graph} to determine software's modules without knowledge of its structure.
    \item Empirically validate our method performance on the \textbf{Linux Kernel Codebase}.

    \item Compare it against state-of-the-art methods (ACDC \cite{acdc}, LIMBO \cite{LIMBO}) and agglomerative clustering methods (agglomerative clustering \cite{average, complete, ward}). 
\end{enumerate}

@@

\item Andritsos and Tzerpos in reference \cite{LIMBO} present an Information-Theoretic 
approach of Software Clustering Algorithms by developing \textbf{LIMBO} which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \textbf{Agglomerative Information Bottleneck} algorithm.
\end{itemize}

@@

\item The edges have the normalized cosine similarities as weights $w(i,j) = (1 + \cos(\mathbf x_i, \mathbf x_j)) / 2$.
\end{enumerate}

\heading{Obtaining a clustering}
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
The Louvain Method is a greedy optimization method that (approximately) maximizes the modularity function of a graph $H(V, E)$

@@

\item it is a large and complex system with 27 years of continuous development. 
\item it spans $\sim$20.3 million lines of source code at the time of writing.
\item  it is easy to establish a ground truth due to its clear structure and construct test-cases for evaluations.

\item it is used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

@@

\item Compare our method against agglomerative clustering baselines (average, complete, and Ward linkage) and state-of-the-art methods (LIMBO and ACDC).
\item Use the MoJo distance metric to calculate the distance between the proposed clustering and ground truth.
\item Evaluate clusterings \textbf{stability}, \textbf{authoritativeness} and \textbf{extremity}.

@@

    \begin{table}
      \centering
      \small
\begin{tabular}{lrrrrrrr}
    \toprule
    \textbf{Algorithm} & \textbf{Feat. Dim}  & $n_c$ & \textbf{Range} & $\bar x$ & $\sigma$ & \textbf{Median} & \textbf{MoJo D.} \\
    \midrule
    \textbf{ACDC}  & -- & 9055 & 1 -- 4245 & 5 & 46 & 2 & 33694\\
    Average Linkage  & \emph{300} & \emph{21} & 1--3406 & 163 & 725 & 1 & 2092 \\
    Complete Linkage  & \emph{300} & \emph{21} & 1--1529 & 163 & 412 & 19 & 1710 \\
    \textbf{LIMBO}  ($B=100, \; S = \infty$) & 12317 &\emph{21} & 50--1810 & 163 & 375 & 50  & 1482 \\

    Ward Linkage & \emph{300} & \emph{21} & 21--948 & 163 & 223 & 70 & 1138 \\

    \textbf{SADE} & \emph{300} & 10 ($\pm$ 2)  & 2 ($\pm$ 0) -132 ($\pm$ 13) & 64 ($\pm$ 4) & 40 ($\pm$ 4) & 65 ($\pm$ 10) & 243 ($\pm$ 1)  \\
    \textbf{SADE} (Directed) & \emph{300} & 5 ($\pm$ 2) & 1 ($\pm$ 1) - 614 ($\pm$ 1) & 141 ($\pm$ 39) & 253 ($\pm$ 25) & 2 ($\pm$ 0.3)  & 237 ($\pm$ 2) \\
    \midrule
    Ground Truth & -- & 21 & 1--1348 & 163 & 341 & 11.0 & -- \\
    \bottomrule
  \end{tabular}

@@

	\begin{itemize}
		\justifying
    \item \textbf{Authoritativeness -- } \textbf{Surpass} all clustering algorithms in terms of MoJo Distance Metric.
    \item \textbf{Extremity -- } Production of \textbf{balanced clusterings}.
    \item \textbf{Stability -- } Production of \textbf{stable clusterings}.
    \item Results were produced \textbf{without knowing the number of clusters of the ground truth a priori}.
    \item Provide a \textbf{simplistic approach} to software clustering combining vector semantics and the call graph.
	\end{itemize}

@@

	\begin{alertblock}{Conclusions}
	\begin{enumerate}
	\justifying
    \item Use of \textbf{vector semantics} and the \textbf{call graph} to produce meaningful clusterings.
    \item Performing our study on a very large system (Linux) gives us \textbf{further insight} on the nature of software itself.
    \item \textbf{Outperform} state-of-the-art and baseline methods in terms of authoritativeness and extremity.
    \item Produce \textbf{stable} and \textbf{balanced} clusterings. 
\end{enumerate}

### 2019-07-04 slides/slides.tex
Add slides with makefile

\begin{enumerate}
    \item First the code is tokenized and identifiers are split into their constituent parts using dynamic programming and n-grams. Stop-words are removed.

    \item For example \texttt{zone\_seqlock\_init} becomes \texttt{zone}, \texttt{seqlock}, \texttt{init} and \texttt{inprogress} becomes \texttt{in} and \texttt{progress}

    \item Let $L(m)$ be the cost of the optimal split in the $m$-th character of the identifier. Then

    $$\small L(m) = 
    \begin{cases}
	0 & m = 0 \\
	\min_{n \le \ell \le m - n} \{ L(\ell) + c(\ell, \ell + n) \} & \text{else}

\end{cases}
$$

	\item The resulting tokens are lemmatized using the English Lemmatizer provided by the spaCy \cite{spacy} package

@@

$$\Pr [w_c \mid w_t] = \frac {\exp(s(w_c, w_t))} {\sum_{j = 1}^{V} \exp(s(w_t, j)}$$

@@

\begin{itemize}
    \item The weights assigned to every edge are the normalized cosine similarities
    $$\cos(\vec x_i, \vec x_j) = \frac {\langle \vec x_i, \vec x_j \rangle} {\| \vec x_i \| \| \vec x_j \|}$$
    $$w(i, j) = \frac {1 + \cos(\vec x_i, \vec x_j)} {2} \qquad \forall (i, j) \in E(H)$$

    \item Experiments were run using both the directed and the undirected version of the graph. The directed version of the graph required doing a bipartite transformation \cite{malliaros} where every edge $(i, j)$ was mapped to $\{ i, j' \}$ where $j'$ was a copy of $j \in V$. After community detection, the communities which $j$ and $j'$ belonged to were merged using a union-find data structure.

@@

\begin{itemize}
    \item The Louvain method for community detection aims to produce communities which maximize the modularity function

@@

    \item \textsc{modularity-optimization} is NP-Hard \cite{modularity}.
    \item The Louvain algorithm progresses agglomeratively by checking for each node $i$ where the largest change in $\Delta Q (i, j) $ occurs for each $j$ of its neighboring communities. 
    \framebreak

    \item The change $\Delta Q(i, j)$ in modularity is derived as

@@

    \item The communities that Louvain Clustering produces

@@

The MoJo \cite{mojo} metric is a clustering distance metric used for comparing software clusterings. The MoJo distance between two clusterings $\mathcal C_1, \mathcal C_2$ is defined as the minimum number of moves and joins to transform one clustering to another where 
\begin{itemize}
    \item \textbf{Move:} Moving a resource from one cluster to another (that includes moving a resource into a previously nonexistent cluster or creating a new cluster.
    \item \textbf{Join:} Join two clusters into one.
\end{itemize}

$$\textrm{MoJo} (\mathcal C_1, \mathcal C_2) = \min \{ \textrm{mno} (\mathcal C_1, \mathcal C_2), \textrm{mno}(\mathcal C_2, \mathcal C_1) \}$$
    Exact computation is not \textbf{efficient} so a \textbf{heuristic} is proposed.
\end{frame}

@@

\begin{itemize}
    \item Average Linkage $\mathrm{argmax}_{A, B} \frac {w(A, B)}{|A||B|}$
    \item Complete Linkage $\mathrm{argmax}_{a \in A, b \in B} w(a, b)$
    \item Single Linkage $\mathrm{argmin}_{a \in A, b \in B} w(a, b)$
    \item Ward Linkage 
    \item Information Loss (Agglomerative Information Bottleneck Algorithm)
\end{itemize}

@@

\begin{frame}{Discussion}

\begin{itemize}
    \item Surpass all clustering algorithms in terms of MoJo Distance Metric
    \item Production of balanced clusterings 
    \item Production of stable clusterings 
    \item Results were produced without knowing the number of clusters of the ground truth a priori
    \item Provide a simplistic approach to software clustering combining vector semantics and the call graph
\end{itemize}

@@

\begin{enumerate}
    \item Use vector semantics and the call graph to produce meaningful clusterings
    \item Performing our study on a very large system (Linux) gives us further insight on the nature of software itself
    \item Outperform state-of-the-art and baseline methods in terms of authoritativeness and extremity
    \item Produce stable and balanced clusterings 
\end{enumerate}

@@

\begin{itemize}
    \item Test our system with various codebases
    \item 
\end{itemize}

>>>>

\begin{enumerate}
    \item<1-> First the code is tokenized and identifiers are split into their constituent parts using dynamic programming and n-grams. Stop-words are removed.

    \item<2-> For example \texttt{zone\_seqlock\_init} becomes \texttt{zone}, \texttt{seqlock}, \texttt{init} and \texttt{inprogress} becomes \texttt{in} and \texttt{progress}

    %\item Let $L(m)$ be the cost of the optimal split in the $m$-th character of the identifier. Then

    %$$\small L(m) = 
    %\begin{cases}
	%0 & m = 0 \\
	%\min_{n \le \ell \le m - n} \{ L(\ell) + c(\ell, \ell + n) \} & \text{else}

%\end{cases}
%$$

	\item<3-> The resulting tokens are lemmatized using the English Lemmatizer provided by the spaCy \cite{spacy} package

@@

$$\Pr [w_c \mid w_t] = \frac {\exp(s(w_c, w_t))} {\sum_{j = 1}^{V} \exp(s(w_t, j))}$$

@@

\begin{itemize}
    \item<1-> The weights assigned to every edge are the normalized cosine similarities
    $$\cos(\vec x_i, \vec x_j) = \frac {\langle \vec x_i, \vec x_j \rangle} {\| \vec x_i \| \| \vec x_j \|}$$
    $$w(i, j) = \frac {1 + \cos(\vec x_i, \vec x_j)} {2} \qquad \forall (i, j) \in E(H)$$

    \item<2-> Experiments were run using both the directed and the undirected version of the graph. The directed version of the graph required doing a bipartite transformation \cite{malliaros} where every edge $(i, j)$ was mapped to $\{ i, j' \}$ where $j'$ was a copy of $j \in V$. After community detection, the communities which $j$ and $j'$ belonged to were merged using a union-find data structure.

@@

\begin{itemize}
    \item<1-> The Louvain method for community detection aims to produce communities which maximize the modularity function

@@

    \item<2-> \textsc{modularity-optimization} is NP-Hard \cite{modularity}.
    \item<3-> The Louvain algorithm progresses agglomeratively by checking for each node $i$ where the largest change in $\Delta Q (i, j) $ occurs for each $j$ of its neighboring communities. 
    \framebreak

    \item<4-> The change $\Delta Q(i, j)$ in modularity is derived as

@@

    \item<5-> The communities that Louvain Clustering produces

@@

The MoJo \cite{mojo} metric is a clustering distance metric used for comparing software clusterings. The MoJo distance between two clusterings $\mathcal C_1, \mathcal C_2$ is defined as the minimum number of moves and joins to transform one clustering to another where 
\begin{itemize}
    \item<1-> \textbf{Move:} Moving a resource from one cluster to another (that includes moving a resource into a previously nonexistent cluster or creating a new cluster.
    \item<2-> \textbf{Join:} Join two clusters into one.
	\item<3-> The MoJo distance is therefore defined as

$$\textrm{MoJo} (\mathcal C_1, \mathcal C_2) = \min \{ \textrm{mno} (\mathcal C_1, \mathcal C_2), \textrm{mno}(\mathcal C_2, \mathcal C_1) \}$$
    \item<4-> Exact computation is not \textbf{efficient} so a \t-extbf{heuristic} is proposed.

@@

\begin{itemize}
    \item Average Linkage\footnote{$w(A, B) = \sum_{a \in A, b \in B} w(a,b)$}  $\mathrm{argmax}_{A, B} \frac {w(A, B)}{|A||B|}$
    \item Complete Linkage $\mathrm{argmax}_{a \in A, b \in B} w(a, b)$
    \item Single Linkage $\mathrm{argmin}_{a \in A, b \in B} w(a, b)$
    \item Ward Linkage 
    \item Information Loss (Agglomerative Information Bottleneck Algorithm)
\end{itemize}

@@

\begin{frame}{Results \& Discussion}

\begin{itemize}
    \item<1-> Surpass all clustering algorithms in terms of MoJo Distance Metric
    \item<2-> Production of balanced clusterings 
    \item<3-> Production of stable clusterings 
    \item<4-> Results were produced without knowing the number of clusters of the ground truth a priori
    \item<5-> Provide a simplistic approach to software clustering combining vector semantics and the call graph
\end{itemize}

@@

\begin{enumerate}
    \item<1-> Use \textbf{vector semantics} and the \textbf{call graph} to produce meaningful clusterings
    \item<2-> Performing our study on a very large system (Linux) gives us further insight on the nature of software itself
    \item<3-> Outperform state-of-the-art and baseline methods in terms of \textbf{authoritativeness} and \textbf{extremity}
    \item<4-> Produce \textbf{stable} and \textbf{balanced} clusterings 
\end{enumerate}

@@

\begin{itemize}
    \item Testing our system with various codebases to validate method's generalizability
    \item Development of evaluation policies with users should be taken into account, especially when dealing with old codebases lacking technical documentation.
    \item Integration with more static analyzers 
\end{itemize}
